Thesis: draft

### 2021-05-21 draft.tex
Update draft.tex
changed Ch.2, erdos-renyi model

Since for large $N$, $\ln{N} \ll N$, we find that the average distance is smaller than the size of the network. Generally, the logarithmic dependence of the average distance $\left<d\right>$ on the network size $N$ is one of the two defining properties of the ``small world phenomenon''; the other property is high clustering \cite{Easley2010}.

@@

This thesis uses three network models: (i) the Erd\H{o}s-R\'{e}nyi model (ER); (ii) the Watts-Strogatz model (WS); and (iii) the stochastic block model, of which two types are implemented: the regular stochastic block model (SBM) and the stochastic Watts-Strogatz block model (SBM-WS).\\
Each of these networks serve specific goals which will be shortly discussed in the following paragraphs.\\
\textbf{The Erd\H{o}s-R\'{e}nyi model}\\
The Erd\H{o}s-R\'{e}nyi model is used as a benchmark model. In this thesis we focus on two network topologies that may impact the formation of echo chambers: a high clustering coefficient and community structure. It is therefore necessary to have a model that lacks both of these features to compare the results with. The Erd\H{o}s-R\'{e}nyi model is one of the most simple random network models that does not display both of these features and is hence the perfect comparison model. \\
It is expected that the random nature of the model and the lack of high clustering and community structure will result in the absence of echo chambers.\\
The Erd\H{o}s-R\'{e}nyi model is also used in the regular stochastic block model to construct the constituent communities.\\
\textbf{The Watts-Strogatz model}\\
The Watts-Strogatz model is used as a general model to investigate the effect of high clustering on the evolution of opinions and the formation of echo chambers.\\
Perra and Rocha showed that a high clustering coefficient enhances the formation of echo chambers \cite{Perra2019}. Here, we will use the Watts-Strogatz model to investigate the interplay between high clustering and the presence of stubborn actors and we will use it as a comparison for networks that posses a community structure without having a high clustering coefficient. We will investigate whether community structure behaves similar to high clustering regarding the appearance of echo chambers.\\
Beside the above, the Watts-Strogatz model is also used to build the constituent communities in the stochastic Watts-Strogatz block model.\\
\textbf{The stochastic block model}\\
One of the main features of the stochastic block model is the ability to generate networks with community structure. We will therefore use this model to investigate the impact of community structure on the evolution of opinions and the formation of echo chambers.\\
It is expected that community structure will enhance the formation of echo chambers compared to the random network case. The question remains, however, whether the influence of community structure on the formation of echo chambers is of the same caliber as that of high clustering. It is also of interest to study the possible relation between the appearance of echo chambers and the modularity of the network.\\
The stochastic Watts-Strogatz block model is used as a model that has both high clustering and community structure. We will check if the combination of both features displays any difference in the formation of echo chambers compared to having only a single feature (either community structure or high clustering).\\
All the above models have a homogeneous degree distribution. We did not focus on network models which generate networks with a heterogeneous degree distribution such as the Barab\'{a}si-Albert model. It might, however, be an interesting extension to investigate such types of models as well.\\
The above models have their added value and benefits. One of their most important plus points is their simplicity. They are the most simple network models that one can create and each of them have characteristics and topologies that can be tuned (clustering for WS, community structure for SBM). This makes it really easy to investigate the effect of these different network topologies without the need to take the interplay with other possible complex network features and topologies into account. This plus side is, however, also one of the possible downsides. Real networks are not simple networks with single, isolated characteristics, but are instead complex networks where many different characteristics and topologies may each play a role. In other words, network models may lack many features of real-world networks. It is thus always tricky, and sometimes even incorrect, to extrapolate the results that are obtained by using network models to real-world situations.\\
The fact that all the above models have a homogeneous degree distribution may also be a drawback, since most real-world networks have a heterogeneous degree distribution.

>>>>

Since for large $N$, $\ln{N} \ll N$, we find that the average distance is smaller than the size of the network. Generally, the logarithmic dependence of the average distance $\left<d\right>$ on the network size $N$ is one of the two defining properties of the ``small world phenomenon''; the other property is high clustering \cite{Easley2010}.\\
\textbf{Erd\H{o}s-R\'{e}nyi model in this thesis}\\
The Erd\H{o}s-R\'{e}nyi model is used as a benchmark model. We focus on two network topologies that may impact the formation of echo chambers: a high clustering coefficient and community structure. It is therefore necessary to have a model that lacks both of these features to compare the results with. The Erd\H{o}s-R\'{e}nyi model is one of the most simple random network models that does not display both of these features and is hence the appropriate comparison model. It is also used in the regular stochastic block model to construct the constituent communities.\\
It is expected that the random nature of the model and the lack of high clustering and community structure will result in the absence of echo chambers.

@@

\textbf{The Watts-Strogatz model}\\
\newline
The Watts-Strogatz model is used as a general model to investigate the effect of high clustering on the evolution of opinions and the formation of echo chambers.\\
\newline
Perra and Rocha showed that a high clustering coefficient enhances the formation of echo chambers \cite{Perra2019}. Here, we will use the Watts-Strogatz model to investigate the interplay between high clustering and the presence of stubborn actors and we will use it as a comparison for networks that posses a community structure without having a high clustering coefficient. We will investigate whether community structure behaves similar to high clustering regarding the appearance of echo chambers.\\
\newline
Beside the above, the Watts-Strogatz model is also used to build the constituent communities in the stochastic Watts-Strogatz block model.\\
\newline
\textbf{The stochastic block model}\\
\newline
One of the main features of the stochastic block model is the ability to generate networks with community structure. We will therefore use this model to investigate the impact of community structure on the evolution of opinions and the formation of echo chambers.\\
\newline
It is expected that community structure will enhance the formation of echo chambers compared to the random network case. The question remains, however, whether the influence of community structure on the formation of echo chambers is of the same caliber as that of high clustering. It is also of interest to study the possible relation between the appearance of echo chambers and the modularity of the network.\\
\newline
The stochastic Watts-Strogatz block model is used as a model that has both high clustering and community structure. We will check if the combination of both features displays any difference in the formation of echo chambers compared to having only a single feature (either community structure or high clustering).\\
\newline
All the above models have a homogeneous degree distribution. We did not focus on network models which generate networks with a heterogeneous degree distribution such as the Barab\'{a}si-Albert model. It might, however, be an interesting extension to investigate such types of models as well.\\
\newline
The above models have their added value and benefits. One of their most important plus points is their simplicity. They are the most simple network models that one can create and each of them have characteristics and topologies that can be tuned (clustering for WS, community structure for SBM). This makes it really easy to investigate the effect of these different network topologies without the need to take the interplay with other possible complex network features and topologies into account. This plus side is, however, also one of the possible downsides. Real networks are not simple networks with single, isolated characteristics, but are instead complex networks where many different characteristics and topologies may each play a role. In other words, network models may lack many features of real-world networks. It is thus always tricky, and sometimes even incorrect, to extrapolate the results that are obtained by using network models to real-world situations.\\
\newline
The fact that all the above models have a homogeneous degree distribution may also be a drawback, since most real-world networks have a heterogeneous degree distribution.
